Dirac neutrino mass generation from Majorana messenger

### 2019-09-11 draft.tex
Updates from Overleaf

\subsection{Fermion dark matter}
Since $\psi_{R1}$ is a singlet under the SM gauge group its thermal relic density is controlled by the Yukawa and $\operatorname{U}(1)_X$ gauge interactions. 
The scenario where $\psi_{R1}$ is self annihilating through $h_{i\alpha}$ interaction resembles the very well known scotogenic model \cite{Ma:2006km} whereas for dominant gauge interactions we have the  Majorana Dark matter with a $Z'$ portal.

%$Z'$ portal~\footnote{For a recent review of Majorana Dark matter with a $Z'$ portal see~\cite{Blanco:2019hah} }
% check Ibarra papers
%also through the annihilation
%$\psi_{1} \psi_{1} \to l _{\alpha}^+ \overline{l}_{\beta}^- $, via
%Yukawa coupling $h^{a}_{i}$. These Yukawa couplings are restricted by
%CLFV and neutrino masses. %But, Yukawa couplings $y^{a}_{j}$ is free.
%In the case where the mass of the other fermions and scalars is nearly
%degenerated to $\psi_{1} $, the processes of coannihilation can be
%relevant for the relic density calculation~\cite{Klasen:2013jpa}.

@@

\subsection{Scalar dark matter}
Contrary to fermionic DM, scalar DM has additional interactions beyond the Yukawa couplings  which can be used to control relic density. This corresponds to the simplified dark matter model of scalar singlet-doblet dark matter, which have been studied in detail in~\cite{Restrepo:2019soi}

@@

Now, if the lightest scalar is mainly a singlet, the DM will be
similar to Singlet Scalar Dark Matter
Model~\cite{Athron:2017kgt}\todo{change for the original ref by Ma et
  al. etc}.
The mass range for the scalar singlet that produces the adequate relic
density is around the Higgs resonance and for masses above $1$TeV. For
this model~\cite{}, the relevant coupling for DM phenomenology is the
Higgs portal trough the Lagrangian

@@

\section{Cosmological constraints}
\label{sec:CosmoConstraints}
A not-so-minimal model can be obtained by introducing additional chiral SM singlet fields. 
In this case we can use the solutions to the anomaly cancellation conditions found for an Abelian gauge symmetry $\operatorname{U(1)_X}$ in~\cite{Campos:2017dgc,Okada:2018tgy}.
We use $f$ ($f$) to denote the $X$-charge of the field $f_R$ ($F_L$) for one generation of the SM. The three linear anomalies in $\operatorname{U(1)_X}$ allows to express three $X$-charges in terms of the other two
\begin{align}
  u=&-e+\frac{2l}{3}\,,& d=& e-\frac{4l}{3}\,,& q=& -\frac{l}{3}\,.
\end{align}
The quadratic anomaly condition is automatically satisfied, while the mixed gauge-gravitational and cubic anomalies depend of any extra singlet quiral fermions of zero hypercharge, like the right-handed counterpart of the Dirac neutrinos. For $\alpha$ extra quiral fields with $X$-charge $n_{\alpha}$, these conditions read
\begin{align}
  \sum_{\alpha} n_{\alpha}+3 (e-2l)=&0\,, &   \sum_{\alpha} n_{\alpha}^3+3 (e-2l)^3=&0\,. &
\end{align}
There are many known solutions for the case in which $e-2l=1$. The full set of standard model charges in terms of $l$ is just
\begin{align}
  u=&-1-\frac{4l}{3}\,,&d=&1+\frac{2l}{3}\,,&q=&-\frac{l}{3}\,,&e=&1+2l\,,&h=&-1-l\,.
  \label{Eq:SMCharges}
\end{align}
The same solution but in terms of $h$ is given in ~\cite{Okada:2018tgy}.

@@

%the charges for the SM particles are in equations~\ref{Eq:SMCharges}. Since these are determined by the anomaly cancellation equations.
%For this operator we have that the charges for the new particles are given by
%\begin{align}
%    \psi=&-\frac{\nu+1}{4}\,,&\eta=&-\frac{\nu+1}{4}-l\,,&\sigma=&\frac{1-3\nu}{4}\,.
%\end{align}
%As we are interested here in the case of the exotic set of X-charges charges:$\mp 4$,$\mp 4$,$\pm 5$, the charges for the model particles are shown in the table~\ref{tab:partcont3}.
%Because the charges of the SM fermions are the same for both operators ($5d$ and $6d$), the $N_{eff}$ restrictions shown in the figure~\ref{fig:Neff} apply to both operators.
%
\begin{table}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|c|}
    \hline  
    Fields     & $\operatorname{SU}(2)_L$ & $\operatorname{U}(1)_Y $ & $\operatorname{U}(1)_{X}$& $\operatorname{U}(1)_{B-L}$& $\operatorname{U}(1)_B$& $\operatorname{U}(1)_D$& $\operatorname{U}(1)_G$\\ \hline
$L $  & $\boldsymbol{2}$& $-1/2$  &  $l$&  $-1$&    $0$ &  $-3/2$&  $-1/2$ \\    
$d_R $  & $\boldsymbol{1}$& $-1/2$  &  $1+2l/3$&  $1/3$&    $1$&  $0$&  $2/3$ \\
$u_R $  & $\boldsymbol{1}$& $+2/3$  &  $-1-4l/3$&  $1/3$&   $-1$&  $1$&  $-1/3$ \\
$Q $  & $\boldsymbol{2}$& $-1/6$  &  $-l/3$& $1/3$&    $0$&  $1/2$&  $1/6$ \\\hline
$e_R $  & $\boldsymbol{1}$& $-1$  &  $1+2l$&  $-1$&    $1$ &  $-2$&  $0$ \\
$H $  & $\boldsymbol{2}$& $-1/2$  &  $-1-l$&  $0$&    $-1$ &  $1/2$&  $-1/2$ \\
$\eta$  & $\boldsymbol{2}$ & $+1/2$  & $3/4-l$& $7/4$& $3/4$ &$9/4$&$5/4$ \\
$S$ & $\boldsymbol{1}$ & $0$  & $3/2$& $3/2$&  $3/2$ & $3/2$& $3/2$ \\
$\sigma$ & $\boldsymbol{1}$ & $0$ & $13/4$& $13/4$&  $13/4$ & $13/4$& $13/4$ \\
\hline
 $\nu_{Ri}$ & $\boldsymbol{1}$ & $0$ & $-4$& $-4$&  $-4$ & $-4$& $-4$\\
$\nu_{R3}$ & $\boldsymbol{1}$ & $0$ & $+5$& $+5$&  $5$ & $5$& $5$\\
$\psi_{R\alpha}$  & $\boldsymbol{1}$ & $0$& $3/4$ & $3/4$&  $3/4$ & $3/4$& $3/4$ \\\hline
$\xi_{L\alpha}$  & $\boldsymbol{1}$ & $0$ & $3/4$& $3/4$ &  $3/4$ & $3/4$& $3/4$\\\hline
  \end{tabular}
  \caption{The new scalars and fermions with their respective charges for the 6-dimensional operator. The SM fields have the usual $U(1)_{B-L}$ assignment. Now $\alpha=1,2$}
    \label{tab:partcont3}
\end{table}
%

@@

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|c|}
    \hline  
    Fields     & $\operatorname{SU}(2)_L$ & $\operatorname{U}(1)_Y $ & $\operatorname{U}(1)_{X}$& $\operatorname{U}(1)_{B-L}$& $\operatorname{U}(1)_B$& $\operatorname{U}(1)_D$& $\operatorname{U}(1)_G$\\ \hline
$L $  & $\boldsymbol{2}$& $-1/2$  &  $l$&  $-1$&    $0$ &  $-3/2$&  $-1/2$ \\    
$d_R $  & $\boldsymbol{1}$& $-1/2$  &  $1+2l/3$&  $1/3$&    $1$&  $0$&  $2/3$ \\
$u_R $  & $\boldsymbol{1}$& $+2/3$  &  $-1-4l/3$&  $1/3$&   $-1$&  $1$&  $-1/3$ \\
$Q $  & $\boldsymbol{2}$& $-1/6$  &  $-l/3$& $1/3$&    $0$&  $1/2$&  $1/6$ \\\hline
$e_R $  & $\boldsymbol{1}$& $-1$  &  $1+2l$&  $-1$&    $1$ &  $-2$&  $0$ \\
$H $  & $\boldsymbol{2}$& $-1/2$  &  $-1-l$&  $0$&    $-1$ &  $1/2$&  $-1/2$ \\
$\eta$  & $\boldsymbol{2}$ & $+1/2$  & $3/2-l$& $5/2$& $3/2$ &$3$&$2$ \\
$S$ & $\boldsymbol{1}$ & $0$  & $3$& $3$&  $3$ & $3$& $3$ \\
$\sigma$ & $\boldsymbol{1}$ & $0$ & $5/2$& $5/2$&  $5/2$ & $5/2$& $5/2$ \\
\hline
 $\nu_{Ri}$ & $\boldsymbol{1}$ & $0$ & $-4$& $-4$&  $-4$ & $-4$& $-4$\\
$\nu_{R3}$ & $\boldsymbol{1}$ & $0$ & $+5$& $+5$&  $5$ & $5$& $5$\\
$\psi_{R\alpha}$  & $\boldsymbol{1}$ & $0$& $3/2$ & $3/2$&  $3/2$ & $3/2$& $3/2$ \\\hline
$\xi_{L\alpha}$  & $\boldsymbol{1}$ & $0$ & $3/2$& $3/2$ &  $3/2$ & $3/2$& $3/2$\\\hline
  \end{tabular}
  \caption{The new scalars and fermions with their respective charges. The SM fields have the usual $U(1)_{B-L}$ assignment. Now $\alpha=1,2$}
    \label{tab:partcont2}
\end{table}

>>>>

%\subsection{Fermion dark matter}
Since $\psi_{R1}$ is a singlet under the SM gauge group its thermal relic density is controlled by the Yukawa and $\operatorname{U}(1)_X$ gauge interactions. 
The scenario where $\psi_{1}$ self-annihilates dominantly through $h_{i\alpha}$-mediated interactions resembles the very well known scotogenic model \cite{Ma:2006km}. 
In this case sizable $h_{i\alpha}$ are required to reproduce the correct DM abundance, which in turn leads to a mild tension with experimental upper bounds on the rates for rare charged lepton decays. 
The phenomenology of this scenario has been extensively studied  \cite{Kubo:2006yx,Sierra:2008wj,Suematsu:2009ww,Gelmini:2009xd,Hambye:2009pw,Aoki:2010tf,Schmidt:2012yg,Ma:2012if,Kashiwase:2012xd,Kashiwase:2013uy,Racker:2013lua,Toma:2013zsa,Molinaro:2014lfa,Vicente:2014wga,Merle:2015gea,Merle:2015ica,Merle:2016scw,Hagedorn:2016dze,Lindner:2016kqk,Hessler:2016kwm}.
On the other hand, when the $Z'$ portal is the main gate to visible sector, $\psi_1$ largely annihilates  into neutrinos (see~\cite{Blanco:2019hah} for a recent review on $Z'$-mediated Majorana DM).
% check Ibarra papers
%also through the annihilation
%$\psi_{1} \psi_{1} \to l _{\alpha}^+ \overline{l}_{\beta}^- $, via
%Yukawa coupling $h^{a}_{i}$. These Yukawa couplings are restricted by
%CLFV and neutrino masses. %But, Yukawa couplings $y^{a}_{j}$ is free.
%In the case where the mass of the other fermions and scalars is nearly
%degenerated to $\psi_{1} $, the processes of coannihilation can be
%relevant for the relic density calculation~\cite{Klasen:2013jpa}.

@@

%\subsection{Scalar dark matter}
Contrary to fermionic DM, scalar DM has additional interactions beyond the Yukawa couplings  which can be used to control relic density. This corresponds to the simplified dark matter model of scalar singlet-doblet dark matter, which have been studied in detail in~\cite{Restrepo:2019soi}

@@

Now, if the lightest scalar is mainly a singlet, the DM will be
similar to Singlet Scalar Dark Matter
Model~\cite{Athron:2017kgt}\todo{change for the original ref by Ma et
  al. etc}.
The mass range for the scalar singlet that produces the adequate relic
density is around the Higgs resonance and for masses above $1$TeV. For
this model~\cite{}, the relevant c

oupling for DM phenomenology is the
Higgs portal trough the Lagrangian

@@

\section{Cosmological constraints}
\label{sec:CosmoConstraints}
A not-so-minimal model can be obtained by introducing additional chiral SM singlet fields. 
In this case we can use the solutions to the anomaly cancellation conditions found for an Abelian gauge symmetry $\operatorname{U(1)_X}$ in~\cite{Campos:2017dgc,Okada:2018tgy,Jana:2019mez}.
We use $f$ ($f$) to denote the $X$-charge of the field $f_R$ ($F_L$) for one generation of the SM. The three linear anomalies in $\operatorname{U(1)_X}$ allows to express three $X$-charges in terms of the other two
\begin{align}
  u=&-e+\frac{2l}{3}\,,& d=& e-\frac{4l}{3}\,,& q=& -\frac{l}{3}\,.
\end{align}
The quadratic anomaly condition is automatically satisfied, while the mixed gauge-gravitational and cubic anomalies depend of any extra singlet quiral fermions of zero hypercharge, like the right-handed counterpart of the Dirac neutrinos. For $\alpha$ extra quiral fields with $X$-charge $n_{\alpha}$, these conditions read
\begin{align}
  \sum_{\alpha} n_{\alpha}+3 (e-2l)=&0\,, &   \sum_{\alpha} n_{\alpha}^3+3 (e-2l)^3=&0\,. &
\end{align}
There are many known solutions for the case in which $e-2l=1$. The full set of standard model charges in terms of $l$ is just
\begin{align}
  u=&-1-\frac{4l}{3}\,,&d=&1+\frac{2l}{3}\,,&q=&-\frac{l}{3}\,,&e=&1+2l\,,&h=&-1-l\,.
  \label{Eq:SMCharges}
\end{align}
The same solution but in terms of $h$ is given in ~\cite{Okada:2018tgy}.

@@

%the charges for the SM particles are in equations~\ref{Eq:SMCharges}. Since these are determined by the anomaly cancellation equations.
%For this operator we have that the charges for the new particles are given by
%\begin{align}
%    \psi=&-\frac{\nu+1}{4}\,,&\eta=&-\frac{\nu+1}{4}-l\,,&\sigma=&\frac{1-3\nu}{4}\,.
%\end{align}
%As we are interested here in the case of the exotic set of X-charges charges:$\mp 4$,$\mp 4$,$\pm 5$, the charges for the model particles are shown in the table~\ref{tab:partcont3}.
%Because the charges of the SM fermions are the same for both operators ($5d$ and $6d$), the $N_{eff}$ restrictions shown in the figure~\ref{fig:Neff} apply to both operators.
%
\begin{table}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|c|}
    \hline  
    Fields     & $\operatorname{SU}(2)_L$ & $\operatorname{U}(1)_Y $ & $\operatorname{U}(1)_{X}$& $\operatorname{U}(1)_{B-L}$& $\operatorname{U}(1)_R$& $\operatorname{U}(1)_D$& $\operatorname{U}(1)_G$\\ \hline
$L $  & $\boldsymbol{2}$& $-1/2$  &  $l$&  $-1$&    $0$ &  $-3/2$&  $-1/2$ \\    
$d_R $  & $\boldsymbol{1}$& $-1/2$  &  $1+2l/3$&  $1/3$&    $1$&  $0$&  $2/3$ \\
$u_R $  & $\boldsymbol{1}$& $+2/3$  &  $-1-4l/3$&  $1/3$&   $-1$&  $1$&  $-1/3$ \\
$Q $  & $\boldsymbol{2}$& $-1/6$  &  $-l/3$& $1/3$&    $0$&  $1/2$&  $1/6$ \\\hline
$e_R $  & $\boldsymbol{1}$& $-1$  &  $1+2l$&  $-1$&    $1$ &  $-2$&  $0$ \\
$H $  & $\boldsymbol{2}$& $1/2$  &  $-1-l$&  $0$&    $-1$ &  $1/2$&  $-1/2$ \\
$\eta$  & $\boldsymbol{2}$ & $+1/2$  & $3/4-l$& $7/4$& $3/4$ &$9/4$&$5/4$ \\
$S$ & $\boldsymbol{1}$ & $0$  & $3/2$& $3/2$&  $3/2$ & $3/2$& $3/2$ \\
$\sigma$ & $\boldsymbol{1}$ & $0$ & $13/4$& $13/4$&  $13/4$ & $13/4$& $13/4$ \\
\hline
 $\nu_{Ri}$ & $\boldsymbol{1}$ & $0$ & $-4$& $-4$&  $-4$ & $-4$& $-4$\\
$\nu_{R3}$ & $\boldsymbol{1}$ & $0$ & $+5$& $+5$&  $5$ & $5$& $5$\\
$\psi_{R\alpha}$  & $\boldsymbol{1}$ & $0$& $3/4$ & $3/4$&  $3/4$ & $3/4$& $3/4$ \\\hline
$\xi_{L\alpha}$  & $\boldsymbol{1}$ & $0$ & $3/4$& $3/4$ &  $3/4$ & $3/4$& $3/4$\\\hline
  \end{tabular}
  \caption{The new scalars and fermions with their respective charges for the 6-dimensional operator. The SM fields have the usual $U(1)_{B-L}$ assignment. Now $\alpha=1,2$}
    \label{tab:partcont3}
\end{table}
%

@@

\begin{table}
  \centering
  \begin{tabular}{|c|c|c|c|c|c|c|c|}
    \hline  
    Fields     & $\operatorname{SU}(2)_L$ & $\operatorname{U}(1)_Y $ & $\operatorname{U}(1)_{X}$& $\operatorname{U}(1)_{B-L}$& $\operatorname{U}(1)_B$& $\operatorname{U}(1)_D$& $\operatorname{U}(1)_G$\\ \hline
$L $  & $\boldsymbol{2}$& $-1/2$  &  $l$&  $-1$&    $0$ &  $-3/2$&  $-1/2$ \\    
$d_R $  & $\boldsymbol{1}$& $-1/2$  &  $1+2l/3$&  $1/3$&    $1$&  $0$&  $2/3$ \\
$u_R $  & $\boldsymbol{1}$& $+2/3$  &  $-1-4l/3$&  $1/3$&   $-1$&  $1$&  $-1/3$ \\
$Q $  & $\boldsymbol{2}$& $-1/6$  &  $-l/3$& $1/3$&    $0$&  $1/2$&  $1/6$ \\\hline
$e_R $  & $\boldsymbol{1}$& $-1$  &  $1+2l$&  $-1$&    $1$ &  $-2$&  $0$ \\
$H $  & $\boldsymbol{2}$& $1/2$  &  $-1-l$&  $0$&    $-1$ &  $1/2$&  $-1/2$ \\
$\eta$  & $\boldsymbol{2}$ & $+1/2$  & $3/2-l$& $5/2$& $3/2$ &$3$&$2$ \\
$S$ & $\boldsymbol{1}$ & $0$  & $3$& $3$&  $3$ & $3$& $3$ \\
$\sigma$ & $\boldsymbol{1}$ & $0$ & $5/2$& $5/2$&  $5/2$ & $5/2$& $5/2$ \\
\hline
 $\nu_{Ri}$ & $\boldsymbol{1}$ & $0$ & $-4$& $-4$&  $-4$ & $-4$& $-4$\\
$\nu_{R3}$ & $\boldsymbol{1}$ & $0$ & $+5$& $+5$&  $5$ & $5$& $5$\\
$\psi_{R\alpha}$  & $\boldsymbol{1}$ & $0$& $3/2$ & $3/2$&  $3/2$ & $3/2$& $3/2$ \\\hline
$\xi_{L\alpha}$  & $\boldsymbol{1}$ & $0$ & $3/2$& $3/2$ &  $3/2$ & $3/2$& $3/2$\\\hline
  \end{tabular}
  \caption{The new scalars and fermions with their respective charges. The SM fields have the usual $U(1)_{B-L}$ assignment. Now $\alpha=1,2$}
    \label{tab:partcont2}
\end{table}
